---
title: "CHP Data Analysis"
author: "Josh Pepper"
date: "March 9, 2016"
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'index.html')) })
output:
  html_document:
    theme: yeti
    toc: true
    toc_float: true
    number_sections: true
---

```{r global_options, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(fig.width=8, echo=TRUE, warning=FALSE, message=FALSE)
```

#Introduction
This file is intended to describe the analysis of the CHP SWITRS data set and display maps and results of the analysis.

#Loading data

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr, verbose = FALSE)
library(ggplot2, verbose = FALSE)
library(leaflet, verbose = FALSE)

switrs_all <- read.csv("Collisions.csv")

#take a subset for only data with lat/long
switrs <- subset(switrs_all, POINT_X!=0 & POINT_Y!=0)

#rename lat and long fields
switrs$lat <- switrs$POINT_Y
switrs$long <- switrs$POINT_X

#add jitter to points
switrs$lat <- jitter(switrs$lat, factor = .8)
switrs$long <- jitter(switrs$long, factor = .8)

#reformat dates
switrs$date <- as.Date(as.character(switrs$DATE_), "%m/%d/%y")
```

##Summary statistics

* Number of rows in the data set: *`r dim(switrs)[1]`*
* Number of columns in the data set: *`r dim(switrs)[2]`*
* Names of columns in the data set:
    + `r names(switrs)`

##First map
Map of all CHP citation in Oakland from 1/1/2013-12/31/2014:

```{r, echo=FALSE}
#map all data from 2013-2014
leaflet() %>% addTiles() %>%
  addCircleMarkers(data=switrs,
                   clusterOptions = markerClusterOptions(spiderfyDistanceMultiplier=2, maxClusterRadius=60),
                   popup=~paste("Date: ", date, "<br/>Time:", TIME_)
  )
```

#Bicyclist data
Notice how one of the columns is titled `BICINJ`. We can map a subset of this data and color code it based on time of accident. I also added a link to more information about each accident using data from the UC Berkeley [Transportation Injury Mapping System (TIPS)](http://tims.berkeley.edu/index.php). TIPS has a [pretty good mapping tool](http://tims.berkeley.edu/tools/gismap/index.php) as well.

```{r, echo=FALSE}
#make bike subset
bike <- subset(switrs, BICINJ==1)

#make color bin
bikeColor <- colorBin(rainbow(25), bike$TIME_, bins=24)


#map all data from 2013-2014
leaflet() %>% addProviderTiles("CartoDB.Positron") %>%
  addCircleMarkers(data=bike, stroke = FALSE, fillOpacity = 0.5, radius=8,
                   popup=~paste("Date: ", date, "<br/>Time: ", TIME_,"<br><a href='http://tims.berkeley.edu/tools/query/collision_details.php?no=",CASEID,"' target='_blank'>Link to Case Details</a>", sep=""),
                   color = ~bikeColor(TIME_)
  ) %>%
      addLegend(title = "Time of collision", pal = bikeColor,
                values = bike$TIME_, opacity = 1, position="bottomleft", labFormat = labelFormat(
                  prefix = '', suffix = ':00',
                  transform = function(x) x/100
                  )
                )
```

